A thermodynamic view of dusty protoplanetary disks: axisymmetric stability

### 2016-11-01 linear.tex
adding kz=0 case

\subsection{Overstable dusty epicycles}
In the special case of $k_z=0$, the incompressible dispersion
relation, Eq. \ref{streaming_incompressible} reduces to 
\begin{align}      %streaming_incompressible
\ii \sigma^2 - \frac{\sigma}{\tepsilon \taus}  - \left(\ii \kappa^2 +
  \frac{k_xF_r}{\epsilon}\right) = 0. \label{kz_zero_si}
\end{align}
If the quadratic term is neglected, then $\imag{\sigma}<0$, implying
stability, as concluded by \cite{youdin05a}.

However, retaining all the terms (but still in the incompressible
limit) can yield instability, as follows. Eq. \ref{kz_zero_si} admit
pure oscillations with $\sigma = -\omega$ provided that 
\begin{align*}
  \omega &= \pm \kappa,\\
  \taus & = \frac{\epsilon\kappa}{\tepsilon k _x\left|F_r\right|}
         \equiv \tau_{\mathrm{s}*}.
\end{align*}
If we linearize about this state of marginal stability by writing
$\sigma = - \kappa + \ii \dd s$ and $\taus = \tau_{\mathrm{s}*} +
\dd\taus$ and assume $|\dd s|\ll \kappa$ then 
\begin{align*} 
\dd s = \frac{2\tepsilon }{1 +
  4\tepsilon^2\tau_{\mathrm{s}*}^2}\dd \taus. 
\end{align*}
Thus given $\tepsilon$ and $k_x$, there are overstable modes with
stopping times slightly larger than $\tau_{\mathrm{s}*}$.

@@

{\bf note: kappa2 accounts for dust effect assuming smallh=0.05}
We solve Eq. \ref{streaming_dispersion} for selected cases where SI  
growth rates have been 
calculated analytically from full two-fluid disk models and measured 
from direct numerical particle-gas simulations 
\citep[namely][]{youdin07b,bai10b}. These authors use normalized
wavenumbers $K_{x,z} = \eta r k_{x,z}$ where
\begin{align} 
  \eta \equiv -\frac{1}{2\rhog r\OmK^2}\frac{\p P}{\p r} = 
  \frac{1}{2\left(1-\tepsilon\right)}\frac{F_r}{r\OmK^2}, 
\end{align} 
measures the pressure offset of Keplerian rotation. We use
$\eta=0.05c_s/r\OmK$. They also define the stopping time as 
$\tau_\mathrm{s}=\tstop/(1-\tepsilon)$.

>>>>

\subsection{Overstable epicycles}
If we further consider the limit $k_z=0$, the incompressible dispersion
relation, Eq. \ref{streaming_incompressible}, reduces to 
\begin{align}      %streaming_incompressible
\ii \sigma^2 - \frac{\sigma}{\tepsilon \taus}  - \left(\ii \kappa^2 +
  \frac{k_xF_r}{\epsilon}\right) = 0. \label{kz_zero_si}
\end{align} 
Here we use $\tau_\mathrm{s}=\tstop/(1-\tepsilon)$ for convenience. 
If the quadratic term is neglected, then $\imag{\sigma}<0$, implying
stability, as concluded by \cite{youdin05a}.

However, retaining all the terms (but still in the incompressible
limit) can yield instability, as follows. Eq. \ref{kz_zero_si} admit
pure oscillations with $\sigma = -\omega$ when 
\begin{align*}
  \omega &= \pm \kappa, \\
  \taus & = \frac{\epsilon\omega}{\tepsilon k _x F_r }
         \equiv \tau_{\mathrm{s}0}.
\end{align*} 
Since $\taus\geq 0$, we require $k_xF_r > 0$ if $\omega = \kappa$;
and $k_xF_r < 0$ if $\omega = -\kappa$. For $k_x>0$ these marginally
stable epicycles require radial pressure gradients decreasing and increasing
outwards, respectively.

%We require $\omega/k_xF_r> 0$

We now perturb about this state of marginal stability by writing
$\sigma = - \omega + \dd \sigma$ with $\dd\sigma = \ii\dd s -
\dd\omega$, $\taus = \tau_{\mathrm{s}0} +
\dd\taus$, and assume the $\dd$ quantities are small in
magnitude. Then 
\begin{align*} 
\dd s = \frac{2\tepsilon\kappa^2 }{1 +
  4\tepsilon^2\tau_{\mathrm{s}0}^2\kappa^2}\dd \taus. 
\end{align*}
Thus given $\tepsilon$ and $k_x$, there exist growing modes
when the stopping time is slightly larger than
$\tau_{\mathrm{s}c}$. These correspond to slowly-growing epicycles
\citep[cf. the `convective overstability', ][]{klahr14,lyra14,latter16}.

@@

{\bf note: kappa2 accounts for dust effect assuming smallh=0.05}

We now consider the streaming instability with $k_z\neq 0$. 
We solve Eq. \ref{streaming_dispersion} for selected cases where SI  
growth rates have been 
calculated analytically from full two-fluid disk models and measured 
from direct numerical particle-gas simulations 
\citep[namely][]{youdin07b,bai10b}. These authors use normalized
wavenumbers $K_{x,z} = \eta r k_{x,z}$ where
\begin{align} 
  \eta \equiv -\frac{1}{2\rhog r\OmK^2}\frac{\p P}{\p r} = 
  \frac{1}{2\left(1-\tepsilon\right)}\frac{F_r}{r\OmK^2}, 
\end{align} 
measures the pressure offset of Keplerian rotation. We use
$\eta=0.05c_s/r\OmK$.
